IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-10-16 paper/intro.tex
s/SWAPI/COWL
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>
Adapting intro

Modern web content is rendered using a potentially large number of
different components with differing provenance.
Disparate and untrusting components may arise from browser
extensions (whose JavaScript code runs alongside website
code), web applications (with possibly untrusted third-party
libraries), and mashups (which combine code and data from
websites that may not even be aware of each other's existence.)
While just-in-time combination of untrusting components
offers great flexibility, it also poses complex security challenges.
In particular, maintaining data privacy in the face of malicious
extensions, libraries, and mashup components has been difficult,
and in general effectively impossible.

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve JavaScript's semantics while retrofitting it with dynamic information
flow control.

One approach to information flow control is fine-grained tracking,
which offers great flexibility.
However, these systems
require deep integration with the programming language and modify
the semantics in non-trivial ways.  This makes fine-grained systems
less suitable in contexts where preserving an existing language
is important.
%
Moreover, the task of adding fine-grained IFC even to a simple language is
non-trivial~\cite{hritcu2013testing}; when considering a real-world language,
such as JavaScript, this challenge becomes almost insurmountable.

In the alternative \textit{coarse-grained} approach to IFC,
a system is divided into relatively coarse computational units,
each with a single label dictating its security policy.
Only
communication between isolated computational units must be tracked.
%A recent system named SWAPI~\cite{swapi} adds IFC to the
%browser by identifying \emph{web workers} as an existing computational
%unit.  Communication between workers is mediated using
%IFC to ensure non-interference, preventing sensitive workers from
%influencing the computation of less sensitive workers.
This coarse-grained approach provides a number of advantages:
(1) minimal changes to an
existing programming language when adding IFC, which (2) allows
the reuse of existing programs.  Furthermore, (3)
reduced runtime overhead because checks need only
be performed at isolation boundaries, compared to fine-grained
systems were the runtime overhead can be impractical~\cite{JSFlow}.
Finally, (4) associating
a single security label with the entire computational unit simplifies
understanding and reasoning about the correctness of the
system, without information flow reasoning about the
technical details of the semantics of the programming language.

In this paper, we take coarse-grained labeling~\cite{Zeldovich:2006,
lio} to its logical extreme: an information flow control system should
be thought of as multiple instances of completely isolated language
runtimes or \emph{tasks}, with information flow control applied to
inter-task communication.  We describe a formal system in which an IFC
system can be designed once and then applied to any programming language
which has control over external effects (e.g., JavaScript or C with
access to hardware privilege separation).  We formalize this system
using an approach by
Matthews-Findler~\cite{Matthews:2007:OSM:1190216.1190220} for combining
operational semantics and prove non-interference guarantees
that are independent of the choice of specific target language.

What distinguishes this setting from previous coarse-grained IFC
systems?
%
First, even though the underlying semantic model involves communicating
tasks, these tasks can be coordinated together in ways that simulate
traditional features in monolithic languages.  In fact, simulating
features in this way is a very useful \emph{design tool} for discovering
what variants of the features are permissible and which are not.
%
Second, although completely separate tasks are semantically easy to
reason about, real-world implementations often blur the lines between
tasks in the name of efficiency.
Characterizing what optimizations are permissible is subtle, since
removing transitions from the operational semantics of a language can
break non-interference.  We partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then non-interference in the abstract specification
carries over to the concrete implementation.

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  dynamic information flow control language free of non-IFC constructs.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Section~\ref{sec:retrofit})
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC language that preserves non-interference. (Section~\ref{sec:formal}
  and~\ref{sec:concrete})
  \item We have implemented an IFC system based on these semantics
  for Node.js, and we connect our formalism to another implementation
  based on this work
  for client-side JavaScript in Gecko~\cite{swapi}.
  Furthermore, we outline an implementation for the C programming
  language.
  (Section~\ref{sec:real})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, labeled references, or a
  notion of clearance. (Section~\ref{sec:extensions})
\end{itemize}

>>>>

Modern web content is rendered using a potentially large number of
different components with differing provenance.
Disparate and untrusting components may arise from browser
extensions (whose JavaScript code runs alongside website
code), web applications (with possibly untrusted third-party
libraries), and mashups (which combine code and data from
websites that may not even be aware of each other's existence.)
%While just-in-time combination of untrusting components
%offers great flexibility, it also poses complex security challenges.
%In particular,
Maintaining data privacy in the face of malicious
extensions, libraries, and mashup components has been rather difficult.
%and in general effectively impossible.

Information flow control (IFC) is a promising technique that provides security
by tracking the flow of sensitive data through a system.  Untrusted code is
confined so that it cannot exfiltrate data, except as per an information flow
policy.  Significant research has been devoted to adding various forms of IFC to
different kinds of programming languages and systems.  In the context of the
web, however, there is a strong motivation to preserve JavaScript's semantics
and avoid any modification to its engine.
%, while retrofitting it with dynamic
%information flow control.

%%% Ale: Related work?
%% One approach to information flow control is fine-grained tracking,
%% which offers great flexibility.
%% However, these systems
%% require deep integration with the programming language and modify
%% the semantics in non-trivial ways.  This makes fine-grained systems
%% less suitable in contexts where preserving an existing language
%% is important.
%
%% Moreover, the task of adding fine-grained IFC even to a simple language is
%% non-trivial~\cite{hritcu2013testing}; when considering a real-world language,
%% such as JavaScript, this challenge becomes almost insurmountable.

Inspired by OS research~\cite{}, we present a \textit{coarse-grained} dynamic
IFC monitor to JavaScript (as specified by ECMAScript~\cite{}). Specifically,
JavaScript code is divided into relatively coarse computational units
(i.e. workers), each with a single label dictating its security policy. As a
result, it is just enough to track communication between isolated computational
units.
%A recent system named COWL~\cite{swapi} adds IFC to the
%browser by identifying \emph{web workers} as an existing computational
%unit.  Communication between workers is mediated using
%IFC to ensure non-interference, preventing sensitive workers from
%influencing the computation of less sensitive workers.
This technique provides a number of advantages: (1) no changes to JavaScript's
semantics when adding IFC ---our approach only exposes an small API to
JavaScript code in order to interact with the monitor---, and
% which (2) allows
%the reuse of existing programs.  Furthermore,
(2) reduced runtime overhead because checks only need to be performed at isolation
boundaries instead of at every instruction in the code~\cite{JSFlow}.
%, compared to fine-grained systems were the runtime overhead can be
% impractical~\cite{JSFlow}.
Finally, (3) associating a single security label with the entire computational
unit simplifies understanding and reasoning about the correctness of the system,
without reasoning about most of the semantics details of JavaScript.
% underlying programming language, i.e., JavaScript.

To reason about our monitor, we introduce a \emph{generic} formalism which
translates coarse-grained IFC ideas~\cite{Zeldovich:2006} into a language-based
setting: an information flow control system should be thought of as multiple
instances of \emph{completely isolated language runtimes} or \emph{tasks}, with
information flow control applied to inter-task communication.
% We describe a
% formal framework in which an IFC system can be designed once and then applied to
% any programming language which has control over external effects like
% JavaScript.
%(e.g., JavaScript) or C with access to hardware privilege separation).
We formalize our framework by using the Matthews-Findler's
approach~\cite{Matthews:2007:OSM:1190216.1190220} for combining operational
semantics and prove non-interference guarantees that are independent of the
choice of the specific underlying programming language, e.g., JavaScript.
%framework to , being one instead of
%that ECMAScript. %target language.

%% There are a number of points that distinguish this setting from
%% previous coarse-grained IFC systems.
%% %
%% First, even though the underlying semantic model involves communicating
%% tasks, these tasks can be coordinated together in ways that simulate
%% features of traditional languages.
%% %
%% In fact, simulating
%% features in this way is a useful \emph{design tool} for discovering
%% what variants of the features are permissible and which are not.
%% %
%% Second, although completely separate tasks are semantically easy to
%% reason about, real-world implementations often blur the lines between
%% tasks in the name of efficiency.
%% Characterizing what optimizations are permissible is subtle, since
%% removing transitions from the operational semantics of a language can
%% break non-interference.  We partially address this issue
%% by characterizing isomorphisms between the operational semantics of our
%% abstract language and a concrete implementation, showing that if this
%% relationship holds, then non-interference in the abstract specification
%% carries over to the concrete implementation.

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained dynamic information
    flow control monitor which is agnostic to non-IFC constructs.
    %We show how a
  \item We show non-interefence results when target languages exhibit a notion
    of isolated computations and control of external side-effects.
    %large class of target languages can be combined with this IFC language and
    %prove that the result provides
    %non-interference. (Sections~\ref{sec:retrofit} and \ref{sec:formal})
  %% \item We provide a proof technique to show the non-interference
  %% of a concrete semantics for a potentially optimized IFC language
  %% by means of an isomorphism and show a class of restrictions on
  %% the IFC language that preserves non-interference. (Sections~\ref{sec:concrete})
  \item
  %We have implemented an IFC system based on these semantics
  %or Node.js, and
  We connect our formalism to COWL~\cite{swapi}, a
  client-side IFC monitor for JavaScript in Firefox.
  %Furthermore, we outline an implementation for the C programming
  %language.
  %(Section~\ref{sec:real})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, labeled references, or a
  notion of clearance. %(Section~\ref{sec:extensions})
\end{itemize}
